---
title: "R Notebook"
output:
  html_document: 
    toc: true
    toc_depth: 4
---

# to do

move this file to the GIT folder.\
Republish the plots

Fix code commenting, etc.

Add to the markdown file

on git, move the md and the sub-folders to a folder with the same name as the md file

## Import packages

```{r}
#| label: "packages"

library(readr)
library(tidyverse)
library(ggplot2)
library(gt)
library(gtExtras)
library(plotly)
library(RColorBrewer)
library(scales)
```

## Import data & Clean

```{r}
#| label: "import data"

trade <- read_csv("trade_2023.csv", skip = 11)
```

```{r}
#| label: "clean_cols"

trade <- janitor::clean_names(trade)

trade <- select(trade, order_date, supplier, recipient, numbers_delivered,
                armament_category, description, designation, status, sipri_estimate,
                tiv_deal_unit, tiv_delivery_values)

# Shorten column names
trade <- trade |>
  rename(
    order_year = order_date,
    arm_cat = armament_category,
    tiv_unit = tiv_deal_unit,
    tiv_delivery = tiv_delivery_values,
    items = numbers_delivered
  )

# Factors
trade$supplier <- factor(trade$supplier, ordered = FALSE)
trade$recipient <- factor(trade$recipient, ordered = FALSE)
trade$arm_cat <- factor(trade$arm_cat, ordered = FALSE)

trade <- trade |>
  filter(
    !(arm_cat %in% c("Engines", "Other", "Satellites"))
  ) |>
  mutate(
    period = case_when(
      order_year %in% c(2006:2013) ~ "P1",
      order_year %in% c(2014:2021) ~ "P2",
      order_year >= 2022 ~ "P3",
      TRUE ~ "P0"
    )
  )

```

## Russia sales - Time Series

```{r}
#| label: "sales_timeseries"

russia_trade_trend <- trade |>
  filter(
    supplier == "Russia",
    order_year > 1991
  ) |>
  group_by(order_year) |>
  summarise(
    TIV = sum(tiv_delivery),
    .groups = "drop"
  ) |>
  ggplot(
    aes(order_year, TIV)
  ) +
 geom_line(
     aes(alpha = .5),
     show.legend = FALSE,
     linewidth = 1
     ) +
  
  geom_point(size = 1.5,
             show.legend = FALSE) + 
  
  geom_vline(xintercept = c(2014, 2022), linetype = 2) +
  
  annotate("text", label = "Crimea",
           angle = 90, hjust = 1, x = 2014.5, y = Inf) +
  
  annotate("text", label = "War",
           angle = 90, hjust = 1, x = 2022.5, y = Inf) +

  theme(axis.text.x = element_text(angle = 0, vjust = 0.5)) +
  
  labs(
    title = "Russiam Arms Exports",
    subtitle = "From 1991 to 2023.",
    y = "TIV",
    x = element_blank(),
    alpha = element_blank(),
    caption = "Stockholm International Peace Research Institute (SIPRI),
    SIPRI Arms Transfers Database, accessed [November, 2023],
    https://www.sipri.org/databases/armstransfers."
    ) +
  theme(plot.title = element_text(face = "bold")) +
  theme_minimal()

russia_trade_trend


russia_trade_trend |> ggsave(filename = "plot_russia_exports_trend.jpg", width = 8, height = 6, dpi = 300)



```

```{r}
#| label: "export_time_series"
#| fig-height: 6
#| fig-width: 8

# total exports
trade |>
  filter(
    supplier == "Russia",
    order_year > 2000
  ) |>
  group_by(order_year) |>
  summarise(
    TIV = sum(tiv_delivery),
    .groups = "drop"
  ) |>
  ggplot(
    aes(order_year,TIV, label = TIV)
  ) +
  geom_col()
 

# total sales, with milestones
russia_trade_cat_trend <- trade |>
  filter(
    supplier == "Russia",
    order_year > 2005
  ) |>
  
  group_by(order_year, arm_cat) |>
  summarise(
    TIV = sum(tiv_delivery),
    .groups = "drop"
  ) |>
  
  group_by(arm_cat) |>
  ggplot(
    aes(order_year,TIV, label = TIV, fill = arm_cat)
  ) +
  
  geom_col(
    show.legend = FALSE
  ) +
  
  geom_vline(xintercept = c(2014, 2022), linetype = 2) +
  annotate("text", label = "Crimea", size = 3,
           angle = 90, hjust = 1, x = 2014.5, y = Inf) +
  annotate("text", label = "War", size = 3,
           angle = 90, hjust = 1, x = 2022.5, y = Inf) +
  
  facet_wrap(~arm_cat, scales = "free_y", ncol = 2) +
  
    labs(
    title = "Russian Arms Exports",
    subtitle = "From 2006 to 2023 by Arms Category.",
    y = "TIV",
    x = element_blank(),
    alpha = element_blank(),
    caption = "Stockholm International Peace Research Institute (SIPRI),
    SIPRI Arms Transfers Database, accessed [November, 2023],
    https://www.sipri.org/databases/armstransfers."
    ) +
  theme_minimal()

russia_trade_cat_trend

russia_trade_cat_trend |>
  ggsave(filename = "plot_russia_exports_category_trend.jpg", width = 8, height = 6, dpi = 300)
```

```{R}
#| label: "compare_cat"

# COMPARE CATEGORIES TIV BEFORE CRIMEA, AFTER CRIMEA

# Create dataset with mean TIV for the periods + change pct

trade_diff <- trade |>
  filter(
    supplier == "Russia",
    order_year >= 2006
  ) |>
  # Total TIV per year
  group_by(period, arm_cat) |>
  mutate(
    TIV = sum(tiv_delivery),
    TIV = case_when(
      period == "P1" ~ round(TIV/8,2),
      period == "P2" ~ round(TIV/8,2),
      period == "P3" ~ round(TIV/2,2),
    )) |>
  summarise(
    TIV = mean(TIV),
    .groups = "drop"
  )

trade_diff_pct <- trade_diff |>
  pivot_wider(names_from = period, values_from = TIV) |>
  mutate(
    dP2_1 = if_else(is.na(P2/P1-1),-1,P2/P1-1),
    dP3_2 = if_else(is.na(P3/P2-1),-1,P3/P2-1),
    dP3_1 = if_else(is.na(P3/P1-1),-1,P3/P1-1)
  ) |>
  mutate(across(P1:dP3_1, ~round(.,2))) |>
  arrange(dP3_1) |>
  # Replace NAs
  mutate(
    across(P1:dP3_1, ~if_else(is.na(.),true = 0,.)
  ))


trade_diff
trade_diff_pct 



# Visualize the totals per period, category

trade_diff_pct |>
  pivot_longer(
    cols = c(P1:dP3_1),
               names_to = "series", values_to = "value"
  ) |>
  filter(
    series %in% c("P1","P2","P3")
  ) |>
  group_by(arm_cat) |>
  ggplot(
    aes(series, value, fill = arm_cat)
  ) +
  geom_col(
    show.legend = FALSE
  ) +
  facet_wrap(~arm_cat, scales = "free_y")


# What categories continued selling?
trade_diff_pct |>
  #select(arm_cat,dP2_1,dP3_2,dP3_1) |>
  arrange(dP2_1)

    trade_diff_pct |>
      select(arm_cat,P1,P2,P3) |>
      arrange(desc(P2))
    
    trade_diff_pct |>
      filter(arm_cat %in% c("Aircraft",
                            "Missiles",
                            "Air defence systems",
                            "Armoured vehicles")) |>
      #select(arm_cat,dP2_1,dP3_2,dP3_1) |>
      arrange(dP2_1)
```

Aircraft, missiles armoured vehicles and ADS were the lifeline after 2014.\
Aircraft actually had a single large order, but remained minuscule after. Armoured vehicles just declined slower but ceased by 2020. Missiles persisted up to 2022 as it declined slowly. Artillery even had some sales up to 2017 but with very low volume.

## What items continued to sell? (categories)

```{r}

# Based on P1 for consistency

cat_list = c("Aircraft",
             "Air defence systems",
             "Armoured vehicles",
             "Missiles",
             "Ships")

```

### PERIOD P1

```{r}
#| label: "sales_P1_items"

exp_P1_cat <- trade |>
  filter(
    period == "P1",
    arm_cat %in% cat_list,
    supplier == "Russia"
  ) |>
  
  group_by(arm_cat, description) |>
  summarise(
    units = sum(items),
    TIV = sum(tiv_delivery)
  ) |>
  arrange(arm_cat, desc(TIV)) |>
  
  group_by(arm_cat) |>
  mutate(
    pct = round(TIV / sum(TIV),2),
    cumsum = cumsum(pct),
    period = "P1"
  )

# items worth >5% of total TIV per category

exp_P1_cat_5pct <- exp_P1_cat |>
  filter(pct > 0.05)


exp_P1_cat_5pct


# Filter designations
exp_P1_desig <- trade |>
  filter(
    period == "P1",
    arm_cat %in% cat_list,
    supplier == "Russia",
    description %in% exp_P1_cat_5pct$description
  ) |>
  
  group_by(arm_cat, description, designation) |>
  summarise(
    units = sum(items),
    TIV = sum(tiv_delivery)
  ) |>
  arrange(arm_cat, desc(TIV)) |>
  
  group_by(arm_cat) |>
  mutate(
    pct = round(TIV / sum(TIV),2),
    cumsum = cumsum(pct),
    period = "P1"
  ) |>
  ungroup()


exp_P1_desig
```

### PERIOD P2

```{r}
#| label: "sales_P2_items"

exp_P2_cat <- trade |>
  filter(
    period == "P2",
    arm_cat %in% cat_list,
    supplier == "Russia"
  ) |>
  
  group_by(arm_cat, description) |>
  summarise(
    units = sum(items),
    TIV = sum(tiv_delivery)
  ) |>
  arrange(arm_cat, desc(TIV)) |>
  
  group_by(arm_cat) |>
  mutate(
    pct = round(TIV / sum(TIV),2),
    cumsum = cumsum(pct),
    period = "P2"
  )

# items worth >5% of total TIV per category

exp_P2_cat_5pct <- exp_P2_cat |>
  filter(pct > 0.05)


exp_P2_cat_5pct



# Filter designations
exp_P2_desig <- trade |>
  filter(
    period == "P2",
    arm_cat %in% cat_list,
    supplier == "Russia",
    description %in% exp_P2_cat_5pct$description
  ) |>
  
  group_by(arm_cat, description, designation) |>
  summarise(
    units = sum(items),
    TIV = sum(tiv_delivery)
  ) |>
  arrange(arm_cat, desc(TIV)) |>
  
  group_by(arm_cat) |>
  mutate(
    pct = round(TIV / sum(TIV),2),
    cumsum = cumsum(pct),
    period = "P2"
  ) |>
  ungroup()


exp_P2_desig

```

### PERIOD P3

```{r}
#| label: "sales_P3_items"

exp_P3_cat <- trade |>
  filter(
    period == "P3",
    arm_cat %in% cat_list,
    supplier == "Russia"
  ) |>
  
  group_by(arm_cat, description) |>
  summarise(
    units = sum(items),
    TIV = sum(tiv_delivery)
  ) |>
  arrange(arm_cat, desc(TIV)) |>
  
  group_by(arm_cat) |>
  mutate(
    pct = round(TIV / sum(TIV),2),
    cumsum = cumsum(pct),
    period = "P3"
  )

# items worth >5% of total TIV per category

exp_P3_cat_5pct <- exp_P3_cat |>
  filter(pct > 0.05)


exp_P3_cat_5pct



# Filter designations
exp_P3_desig <- trade |>
  filter(
    period == "P3",
    arm_cat %in% cat_list,
    supplier == "Russia",
    description %in% exp_P3_cat_5pct$description
  ) |>
  
  group_by(arm_cat, description, designation) |>
  summarise(
    units = sum(items),
    TIV = sum(tiv_delivery)
  ) |>
  arrange(arm_cat, desc(TIV)) |>
  
  group_by(arm_cat) |>
  mutate(
    pct = round(TIV / sum(TIV),2),
    cumsum = cumsum(pct),
    period = "P3"
  ) |>
  ungroup()


exp_P3_desig
```

```{r}

# Bind tables from PERIOD P1, P2, P3 above
exp_table <- rbind(exp_P1_cat_5pct,
                   exp_P2_cat_5pct,
                   exp_P3_cat_5pct)

exp_table2 <- exp_table |>
  relocate(period) |>
  select(-pct, -cumsum) |>
  pivot_wider(names_from = period, values_from = c(units, TIV)) |>
  arrange(arm_cat,desc(TIV_P1)) |>
  mutate(
    across(units_P1:TIV_P3, ~if_else(is.na(.), 0,round(.)))
  )

cols_labels <- unlist(lapply(c("Units", "Mean Yearly TIV"), function(x) paste(x, c("P1", "P2", "P3*"), sep = "_"))) 

cols_names <- c("arm_cat", "description") |> append(cols_labels)
names(exp_table2) <- cols_names
  

exp_table2


# Russian exports per category in P1, P2 ,P3

table_russia_exports <- exp_table2 |>
  gt(row_group_as_column = TRUE ) |>
  opt_row_striping() |>
  tab_spanner_delim(delim = "_") |>
  cols_width(
    starts_with("Units") ~ px(50),
    starts_with("Mean") ~ px(100)
    ) |>
  
  tab_header(
    title = "Russian Arms Exports",
    subtitle = "Analysis of Pre-Crimea, Post-Crimea, and Post-Full Scale War Periods"
    ) |>
  
  tab_footnote(
    footnote = md("(*) P3 averages two years (2022-23),
                  compared to eight years in P1 (2006-13) and P2 (2014-21).")
    ) |>
  tab_footnote(
    footnote = md("Source: Stockholm International Peace
    Research Institute (SIPRI),
    SIPRI Arms Transfers Database.")
    ) |>
  
  tab_style(
      style = list(cell_text(weight = "bold")),
      locations = list(cells_row_groups(), cells_column_labels(), cells_title()))


table_russia_exports


gtsave(table_russia_exports, "table_russia_exports_P1_P2_P3.png")

```

## What items continued to sell? (items)

```{r}

# Table with designation breakdown

exp_table_desig <- rbind(exp_P1_desig, exp_P2_desig, exp_P3_desig)

exp_table_desig2 <- exp_table_desig |>
  relocate(period) |>
  select(-pct, -cumsum) |>
  pivot_wider(names_from = period, values_from = c(units, TIV)) |>
  arrange(arm_cat,desc(TIV_P1)) |>
  mutate(
    across(units_P1:TIV_P3, ~if_else(is.na(.), 0,round(.)))
  )

cols_names2 <- c("arm_cat", "description", "Designation") |>
  append(cols_labels)

names(exp_table_desig2) <- cols_names2
  

exp_table_desig2

  
```

```{r}
# Split the data by category
split_data <- exp_table_desig2 |>
  arrange(arm_cat, description, desc(`Mean Yearly TIV_P1`)) |>
  group_split(arm_cat)


# GT table for each category
exp_tables_desig <- lapply(split_data, function(group) {
  
  # Category labels
  arm_cat_label <- unique(group$arm_cat)

  group |>
    select(-arm_cat) |>
    group_by(description) |>
    gt(row_group_as_column = TRUE) |>
    opt_row_striping() |>
    tab_spanner_delim(delim = "_") |>
    cols_width(
      starts_with("Units") ~ px(50),
      starts_with("Mean") ~ px(100)
      ) |>
    
  # Custom title for each table
  tab_header(
    title = paste0(arm_cat_label, " - Russian Arms Exports"),
    subtitle = "Analysis of Pre-Crimea, Post-Crimea, and Post-Full Scale War Periods"
    ) |>
  
  tab_footnote(
    footnote = md("(*) P3 averages two years (2022-23),
                  compared to eight years in P1 (2006-13) and P2 (2014-21).")
    ) |>
  tab_footnote(
    footnote = md("Source: Stockholm International Peace
    Research Institute (SIPRI),
    SIPRI Arms Transfers Database.")
    ) |>
  
  tab_style(
      style = list(cell_text(weight = "bold")),
      locations = list(cells_row_groups(), cells_column_labels(), cells_title())
      )
})


exp_tables_desig


for (i in seq_along(exp_tables_desig)) {
  gtsave(exp_tables_desig[[i]],
         filename = paste0("table_exports_russia_",
                          snakecase::to_snake_case(sub("\\ -.*", "", tolower(unclass(exp_tables_desig[[i]])$`_heading`$title))),
                           ".png"))
}


```

## Buyer Countries and their trade in P1, P2

```{r}
#| label: "countries_pareto"

# List of countries supplied
ctr_ru_cat <- trade |>
  filter(
    #arm_cat %in% cat_list,
    supplier == "Russia",
    period %in% c("P1","P2","P3")
  ) |>
  group_by(period, recipient, arm_cat) |>
  summarise(
    orders = n(),
    units = sum(items),
    TIV = sum(tiv_delivery)
    ) |>
  # arrange(period, desc(TIV)) |>
  # group_by(period) |>
  # mutate(
  #   pct = TIV / sum(TIV),
  #   cumsum = cumsum(pct)
  # ) |> 
  ungroup()


# Count of countries supplied per period
ctr_ru_cat |>
  group_by(period, recipient) |>
  summarise(
    TIV = sum(TIV)
  ) |>
  arrange(desc(TIV)) |>
  summarise(
    n = n()
  )


# Pareto - cumsum of pct
ctr_ru_cat |>
  group_by(period, recipient) |>
  summarise(
    TIV = sum(TIV)
  ) |>
  arrange(desc(TIV)) |>

  group_by(period) |>
  mutate(
    pct = TIV / sum(TIV),
    cumsum = cumsum(pct),
    rank = 1:n()
  ) |>
  relocate(rank) |>
  
  ggplot(
    aes(rank, cumsum, color = period)
  ) +
  geom_line(
    linewidth = 1
  )



  


```

```{r}
#| label: "countries_list"

# Indicator columns for presence of sales
ctr_ru_mean <- ctr_ru_cat |>
  group_by(period, recipient) |>
  summarise(
    TIV = sum(TIV),
    .groups = "drop"
  ) |>
  select(period, recipient, TIV) |>
  pivot_wider(names_from = period, values_from = TIV) |>
  mutate(
    ID_P1 = if_else(is.na(P1), 0, 1),
    ID_P2 = if_else(is.na(P2), 0, 1), # 0 if Px has zero sales
    ID_P3 = if_else(is.na(P3), 0, 1),
    P1 = P1/8,
    P2 = P2/8,
    P3 = P3/2
  )


ctr_ru_mean_group <- ctr_ru_mean |>
  mutate(
    group = case_when(
      (ID_P1 == 1 & ID_P2 == 0 & ID_P3 == 0) ~ "A", # no sales after P1
      (ID_P1 == 1 & ID_P2 == 1 & ID_P3 == 0) ~ "B", # no sales after P2
      (ID_P1 == 0 & (ID_P2 == 1 | ID_P3 == 1)) ~ "C", # no sales BEFORE p1
      (ID_P1 == 1 & ID_P2 == 1 & ID_P3 == 1) ~ "D", # persistent buyers
      TRUE ~ "-"
    ),
    dP2_1 = case_when(
      (is.na(P1) & is.na(P2)) ~ 0,
      (is.na(P1) & !is.na(P2)) ~ P2,
      (!is.na(P1) & is.na(P2)) ~ -P1,
      (!is.na(P1) & !is.na(P2)) ~ P2-P1
    ),
    dP3_2 = case_when(
      (is.na(P2) & is.na(P3)) ~ 0,
      (is.na(P2) & !is.na(P3)) ~ P3,
      (!is.na(P2) & is.na(P3)) ~ -P2,
      (!is.na(P2) & !is.na(P3)) ~ P3-P2
    )
  )
  
ctr_ru_mean_group


ctr_ru # TOTAL TIV per country/category
ctr_ru_total # YEARLY TIV AVERAGE of Px per country
ctr_ru_total_group # grouped by indicators + absolute differences

```

We know which countries stopped buying. Now we need to find out how their volume of sales changed in the same period.

```{r}

# Join data from overall orders to orders to Russia
ctr_exp_compare_all_ru <- left_join(
  ctr_all_mean, 
  ctr_ru_mean_group |>
    select(recipient, group, P1, P2, P3, dP2_1, dP3_2),
  by = join_by(recipient),
  suffix = c("_ALL", "_RU")
  ) |>
  mutate(across(c(P1_ALL: dP3_2_ALL, P1_RU:dP3_2_RU), ~round(.,2)))


# Calculate percentual changes between periods, add indicators,
ctr_exp_compare_all_ru <- ctr_exp_compare_all_ru |>
  mutate(
    across(c(P1_ALL: dP3_2_ALL, P1_RU:dP3_2_RU), ~if_else(is.na(.),0,.))) |>
  mutate(
    P2_1_change_total = round(P2_ALL / P1_ALL -1, 2),
    P2_1_change_ru = round(P2_RU / P1_RU -1, 2),
    P3_2_change_total = round(P3_ALL / P2_ALL -1, 2),
    P3_2_change_ru = round(P3_RU / P2_RU -1, 2),
    russia_lost = as_factor(if_else(P2_1_change_ru < 0, 1, 0))
  )

```

```{r}
#| fig-height: 11
#| fig-width: 8


# Transformation to display outliers (cubic root)
signed_sqrt_trans <- trans_new(
  "signed_sqrt",
  transform = function(x) if_else(x > 1, x^(1/3), x), 
  inverse = function(x) if_else(x > 1, x^3, x)
)

# Labels
group_name <- c(
  "1" = "Q1",
  "2" = "Q2",
  "3" = "Q3",
  "4" = "TIV from Russia - 4th Quartile"
)


# Get quartile ranges
quartile_range_table <- ctr_exp_compare_all_ru |>
  select(
    recipient, group, P1_RU, russia_lost,
         P2_1_change_total, P3_2_change_total,
         P2_1_change_ru, P3_2_change_ru
  ) |>
  filter(
    group %in% c("A", "B", "D"), # Excludes new clients
  ) |>
  mutate(quartile = factor(ntile(P1_RU, 4),
                           levels = c("4", "3", "2", "1"),
                           ordered = TRUE)) |>
  arrange(desc(P1_RU)) |>
  group_by(quartile) |>
  summarise(
    range = round(range(P1_RU), 1)
  )

quartile_range <- ""
for (i in 1:4){
  quartile_range <- paste0(quartile_range,
      "Q",
      i,
      ": From ",
        filter(slice(quartile_range_table,1), quartile == i) |> pull(range),
      " to ",
        filter(slice(quartile_range_table,2), quartile == i) |> pull(range),
      " TIV",
      "\n")
}


# percentual reduction in mean TIV for each period, for all purchases and for Russia

plot_country_trade_russia <- ctr_exp_compare_all_ru |>
  select(
    recipient, group, P1_RU, russia_lost,
    P2_1_change_total, P3_2_change_total,
    P2_1_change_ru, P3_2_change_ru
  ) |>
  filter(
    group %in% c("A", "B", "D"), # Excludes new clients
  ) |>
  mutate(quartile = factor(ntile(P1_RU, 4),
                           levels = c("4", "3", "2", "1"),
                           ordered = TRUE)) |>
  ggplot() +
  
  geom_segment(
    aes(x = 0, xend = P2_1_change_ru, y = reorder(recipient, P2_1_change_total)),
    color="black",
    linewidth = 1
    ) +
  
  geom_col(
    aes(P2_1_change_total, recipient),
    fill = "grey"
  ) +
  
  geom_point(
    aes(P2_1_change_ru, recipient, color = russia_lost),
    size = 2.5,
    show.legend = FALSE
  ) +
  
  facet_grid(
    quartile~., scales = "free_y", space = "free",
    labeller = labeller(quartile = group_name), switch = "y" 
    ) +
  
  labs(
    title = "Recipient's change in overall orders (bar) vs. orders to Russia (dot)",
    subtitle = "Percentual change in mean yearly TIV between 2014-2021 and 2006-2013.",
    y = element_blank(),
    x = "Percentual Change",
    color = element_blank(),
    caption = "Stockholm International Peace Research Institute (SIPRI),
    SIPRI Arms Transfers Database, accessed [November, 2023],
    https://www.sipri.org/databases/armstransfers."
    ) +
  
  theme_minimal() +
  theme(
    plot.title = element_text(face = "bold"),
    panel.grid.minor.x = element_blank(),
    strip.text.y = element_text(face = "bold"),
    strip.placement = "outside",
    strip.background = element_rect(fill = "grey90", color = NA)
    ) +
  
  scale_x_continuous(
    trans = signed_sqrt_trans,
    breaks = as.integer(c(-1, -0.5, 0, 0.5, 1:5, 10, 100 ))
  ) +
  scale_color_brewer(palette = "Set1", direction = -1)


plot_country_trade_russia


plot_country_trade_russia |>
  ggsave(filename = "plot_buyers_trade_overall_russia.jpg", width = 8, height = 11, dpi = 300)
  

# How many countries reduced / not reduced orders to russia?

ctr_exp_compare_all_ru |>
  select(
    recipient, group, russia_lost,
    P2_1_change_total, P3_2_change_total,
    P2_1_change_ru, P3_2_change_ru
  ) |>
  filter(
    group %in% c("A", "B", "D"), # Excludes new clients
  ) |>
  pull(russia_lost) |>
  table()

ctr_exp_compare_all_ru |>
  filter(
    group %in% c("A", "B", "D"), # Excludes new clients
  ) |>
  select(recipient, P3_RU) |>
  filter(P3_RU > 0)
  
  

```

```{r}

# percentual change in total TIV

buyers_imports <- ctr_exp_compare_all_ru |>
  select(
    group, russia_lost,
    P1_ALL, P2_ALL, P3_ALL, P1_RU, P2_RU, P3_RU,
    P2_1_change_total, P2_1_change_ru,
    P3_2_change_total, P3_2_change_ru
  ) |>
  filter(
    group %in% c("A", "B", "D"), # Excludes new clients
  ) |>
  summarise(
    P2_ALL_CHANGE = sum(P2_ALL)/sum(P1_ALL)-1,
    P2_RU_CHANGE = sum(P2_RU)/sum(P1_RU)-1,
    P3_ALL_CHANGE = sum(P3_ALL)/sum(P1_ALL)-1,
    P3_RU_CHANGE = sum(P3_RU)/sum(P1_RU)-1
  ) |>
  mutate(across(c(P2_ALL_CHANGE:P3_RU_CHANGE), ~round(., 3))) |>
  
  rename(
    'Post-Crimea_Overall' = P2_ALL_CHANGE,
    'Post-Crimea_Russia' = P2_RU_CHANGE,
    'Post-War*_Overall' = P3_ALL_CHANGE,
    'Post-War*_Russia' = P3_RU_CHANGE  
  ) |>
  gt() |>
  
  tab_spanner_delim(delim = "_") |>
  fmt_percent(
      decimals = 1
    ) |>
  cols_align("center") |>
  cols_width(
    everything() ~ px(100)
    ) |>
  
  tab_header(
    title = "Change in Traded Volume\nby Russian Arms Importer"
    ) |>
  
  tab_footnote(
    footnote = md("(*) Post-War averages two years (2022-23),
                  compared to eight years pre (2006-13).")
    ) |>
  tab_footnote(
    footnote = md("Source: Stockholm International Peace
    Research Institute (SIPRI),
    SIPRI Arms Transfers Database.")
    ) |>
  
  tab_style(
      style = list(cell_text(weight = "bold", size = 15)),
      locations = list(cells_title())
      )


buyers_imports |> gtsave("table_buyers_imports_change.png")
```

```{r}

russia_lost_list <- ctr_exp_compare_all_ru |>
  filter(russia_lost == 1) |>
  pull(recipient)


# Sum of TIV in P1 and P2 for countries where Russia lost market share
ctr_ru_loss_tiv <- trade |>
  filter(
    recipient %in% russia_lost_list,
    period %in% c("P1","P2")
  ) |>
  group_by(period, recipient, supplier) |>
  summarise(
    TIV = sum(tiv_delivery),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = period, values_from = TIV) |>
  mutate(
    across(c("P1","P2"), ~if_else(is.na(.),0,.)),
    diff = P2-P1
    )

# Top 5 countries represent ~78% of
# the Russian total TIV loss between P1 and P2 
ctr_ru_loss_tiv_top5_list <- ctr_ru_loss_tiv |>
  arrange(diff) |>
  filter(
    supplier == "Russia"
  ) |>
  top_n(-5) |>
  pull(recipient)
  


# suppliers who gained share in coutries which russia lost share
ctr_ru_loss_tiv_competitors <- ctr_ru_loss_tiv |>
  filter(
    recipient %in% ctr_ru_loss_tiv_top5_list
  ) |>
  arrange(desc(diff))

# Total TIV increase of suppliers who
# gained share in coutries which russia lost share
ctr_ru_loss_tiv_competitors |>
  group_by(supplier) |>
  summarise(
    TIV_increase = sum(diff)
  ) |>
  arrange(desc(TIV_increase))


```

```{r}

# Sum of TIV PER CATEGORY in P1 and P2 for the
# top 5 countries where Russia lost market share
ctr_ru_loss_tiv_cat <- trade |>
  filter(
    period %in% c("P1","P2"),
    recipient %in% ctr_ru_loss_tiv_top5_list
  ) |>
  group_by(period, recipient, supplier, arm_cat) |>
  summarise(
    TIV = sum(tiv_delivery),
    .groups = "drop"
  ) |>
  pivot_wider(names_from = period, values_from = TIV) |>
  mutate(across(c("P1","P2"), ~if_else(is.na(.),0,.))) |>
  arrange(desc(P1)) |>
  mutate(
  diff = P2-P1
  )
  
# This answers where the demand has moved to.
# The top 5 countries who cut orders from Russia
# have increased their orders to the following suppliers

ctr_ru_loss_tiv_cat_increase <- ctr_ru_loss_tiv_cat |>
  filter(
    supplier != "Russia"
  ) |>
  group_by(supplier, arm_cat) |>
  summarise(
    TIV_increase = sum(diff)
  ) |>
  arrange(desc(TIV_increase))

ctr_ru_loss_tiv_cat_increase


ctr_ru_loss_tiv_cat_increase_cat_list <- ctr_ru_loss_tiv_cat_increase |>
  pull(arm_cat) |>
  unique()

ctr_ru_loss_tiv_cat_increase_ctr_list <- ctr_ru_loss_tiv_cat_increase |>
  pull(supplier) |>
  unique()


```

```{r}

# What items have the top 5 sellers to the countries which Russia lost share?

# ctr_ru_loss_tiv_cat_increase_cat_list
# ctr_ru_loss_tiv_cat_increase_ctr_list 
# ctr_ru_loss_tiv_top5_list

gainers_cat_items <- trade |>
  filter(
    period %in% c("P1","P2"),
    recipient %in% ctr_ru_loss_tiv_top5_list,
    supplier %in% ctr_ru_loss_tiv_cat_increase_ctr_list,
    arm_cat %in% ctr_ru_loss_tiv_cat_increase_cat_list
  ) |>
  group_by(period, supplier, recipient, arm_cat, designation) |>
  summarise(
    items = sum(items),
    TIV = sum(tiv_delivery),
    .groups = "drop"
  )
  

gainers_cat_items |> arrange(desc(TIV)) # Top TIV
gainers_cat_items |> arrange(desc(items)) # Top Items count


#### Difference between Pre-Crimea and Post-Crimea in the number of items and TIV for each item in the selected categories
#### supplied by the countries who gained share where Russia lost share

### Review this to get insights
### Check the production date as some may not be available before 2014

gainers_cat_items_diff <- gainers_cat_items |>
  pivot_wider(
    names_from = period, values_from = c(items, TIV)
  ) |>
  mutate(
    across(c(items_P1:TIV_P2), ~if_else(is.na(.), 0, .)),
    diff_items = items_P2 - items_P1,
    diff_TIV = TIV_P2 - TIV_P1,    
    ) |>
  filter(
    diff_TIV > 0,
    #items_P1 == 0
    ) |>
  arrange(desc(diff_TIV))


### Grouped tables
gainers_cat_items_diff_table_group <- gainers_cat_items_diff |>
  select(supplier, arm_cat, designation, recipient, items_P1, diff_items, diff_TIV) |>
  arrange(arm_cat, desc(diff_TIV)) |>
  group_split(recipient)
  


gainers_cat_items_diff_table_group_gt <- lapply(gainers_cat_items_diff_table_group, function(group) {
  
  # Category labels
  recipient_label <- unique(group$recipient)

  group |>
    select(-recipient) |>
    rename(
      `Pre-Crimea_Items` = items_P1,
      `Post-Crimea_Items` = diff_items,
      `Post-Crimea_TIV` = diff_TIV,
      Supplier = supplier,
      Designation = designation
    ) |>
    mutate(`Post-Crimea_TIV` = as.integer(`Post-Crimea_TIV`)) |>
    group_by(arm_cat) |>
    gt(row_group_as_column = TRUE) |>
    opt_row_striping() |>
    tab_spanner_delim(delim = "_") |>
    cols_width(
      cols_names = c(`Post-Crimea_Items`, `Post-Crimea_TIV`) ~ px(75)
      ) |>
    cols_width(
      cols_names = c(`Pre-Crimea_Items`) ~ px(120)
      ) |>
    
  # Custom title for each table
  tab_header(
    title = paste0(recipient_label, " -  Shifts in Arms Trade"),
    subtitle = md("Sales by Countries Filling Russia's Market Gap")
    ) |>
  
  tab_footnote(
    footnote = md("Source: Stockholm International Peace
    Research Institute (SIPRI),
    SIPRI Arms Transfers Database.")
    ) |>
  
  tab_style(
      style = list(cell_text(weight = "bold")),
      locations = list(cells_row_groups(), cells_column_labels(), cells_title())
      ) |> 
    cols_align("left", columns = c(Supplier, Designation)) |>
    cols_align("center", columns = c(`Pre-Crimea_Items`, `Post-Crimea_Items`))
})


for (i in seq_along(gainers_cat_items_diff_table_group_gt)) {
  gtsave(gainers_cat_items_diff_table_group_gt[[i]],
         filename = paste0("table_exports_winners_",
           sub("\\ -.*", "",
           unclass(
             gainers_cat_items_diff_table_group_gt[[i]]
             )$`_heading`$title),
             ".png"))
}


# List of designations to check the there are any sales before 2013
# looking for items that were not available in P1 at all

design_list <- gainers_cat_items_diff$designation

trade |>
  filter(
    designation %in% design_list,
    period %in% c("P1","P2")
    ) |>
  group_by(period, designation) |>
  summarise(
    total = sum(items)
  ) |>
  pivot_wider(names_from = period, values_from = total) |>
  filter(
    P1 == 0
  )

#### ALL ITEMS HAVE SOME SALE IN P1. 

```
